Introduction to R - 9/7/17

Vectors
```{r}

a <- c(15,12,10,93) #c means make a vector of these things
a[2] #calls a certain location, three different ways
a[c(2,4)]
a[c(FALSE, TRUE,TRUE,FALSE)]
a*5
a + c(1,1,10,-50)
a + c(1,5) #it repeats the these commands across a i.e. +1 +5 +1 +5

```
Data Frames
- can also be accessed by dataframe[row,column]

```{r}

hypoxiaTested[,1]
hypoxiaTested[1:20,1]

```

Functions
- things you can do with variables

plot()
ttest()
read.table()
- to give R a file path, starting with a / is the absolute path from the computer root
- just a file name suggests its in THE working directory
- ../filename means go up one, and then look
- header(true/false) to let it know if you do/do not have a header
- dec specifies what character represents decimals [this also tells R that your data is numeric]
- if you want to save it (which you do), give it a name <- read.table(hi)

use help, ?, google, tabbing to figure out how to full these functions

use $ to access a column within a data frame
also use $ to ADD a column to an existing data frame

Visualizing Data

histograms! scatterplots! hist() and plot()

Do a Simple T-Test with random normal data
- significance is based on difference in means and n when doing a t test
- major assumption of t-test is equal standard deviations//variances
```{r}
set.seed(seed = 400)
meanzero <- rnorm(20, 0, 1) #100 values, mean zero, SD 1
meanplus <- rnorm(20, 0.01, 1)

hist(meanplus)
hist(meanzero, add = T, col = "red")

t.test(meanzero, meanplus)
```
